---
title: "Exploratory Data Analysis"
format: html
editor: visual
---

### Libraries

```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(moments)
```

### Import Data

```{r}
# import data
# avocado_data <- read_csv("data/avocado.csv") 
# avocado_data 
# 
# avocado_data <- avocado_data %>% select(-1)
# 
# avocado_data <- avocado_data %>% clean_names()


avocado_data <- read_csv("data/avocado.csv") %>%
  select(-1) %>% clean_names()

```

### Understanding the Structure of the Data

```{r}
str(avocado_data)
```

### Quick glimpse of data

```{r}
glimpse(avocado_data)
```

### Quick Summary Statistics

```{r}
summary(avocado_data)
```

### Skim a data frame, getting useful summary statistics

```{r}
skim(avocado_data)
```

### Missing Values

```{r}
colSums(is.na(avocado_data))
sum(is.na(avocado_data))
```

### Distribution Plot for Numerical values

**Density plot**

```{r}
ggplot(avocado_data, aes(x = average_price)) +
  geom_density(fill = "#4a7337", alpha = 0.5) +
  ggtitle("Distribution Price") +
  theme_minimal()
```

**Boxplot**

```{r}
ggplot(avocado_data, 
       aes(x = average_price)) +
  geom_boxplot(fill= "#704012") +
  ggtitle("Distribution Price") +
  theme_minimal()
```

**Histogram**

```{r}
ggplot(avocado_data, aes(x = average_price)) +
  geom_histogram(fill = "#4a7337", color="white", alpha = 0.5) +
  ggtitle("Distribution Price") +
  theme_minimal()

```

Showing the mean and median in a distribution plot

```{r}
ggplot(avocado_data, aes(x = average_price)) +
  geom_histogram(fill = "#4a7337", bins = 30, alpha = 0.5) +
  geom_vline(aes(xintercept = mean(average_price)), 
             color = "blue", linetype = "dashed", size = 1) +  # Mean line
  geom_vline(aes(xintercept = median(average_price)), 
             color = "red", linetype = "dashed", size = 1) +   # Median line
  ggtitle("Distribution of Price") +
  theme_minimal()
```

### **Categorical Data Analysis**

-   Year

```{r}
avocado_data %>% 
  ggplot(aes(year))+
  geom_bar()
```

-   Avocado Types

```{r}
avocado_data %>% 
  ggplot(aes(type))+
  geom_bar()

```

### Bivariate data analysis

```{r}
ggplot(avocado_data, aes(x = average_price, 
                         y = type)) +
  geom_boxplot(fill = c("#704012", "#4a7337")) +
  theme_minimal()
```

-   Average Price vs Year

```{r}
ggplot(avocado_data, aes(x = average_price, 
                         y = factor(
                           as.character(year)
                           )
                         )
       ) +
  geom_boxplot(fill = c("#704012", "#4a7337", "#6b8c21", "#ddd48f")) +
  theme_minimal()+
  labs(x="Average Price", y="Year")
```

-   Average Price vs Date

```{r}
avocado_data %>% 
  ggplot(aes(x = date, y = average_price)) + 
  geom_line(color = "#012f49") +  
  labs(title = "AveragePrice vs Date", x = "Date", y = "Average Price") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  
```

-   *Break dates down into months*

```{r}

avocado_data %>% 
  ggplot(aes(x = date, y = average_price)) + 
  geom_line(color = "#012f49") +  # You can adjust the color as needed
  labs(title = "AveragePrice vs Date", x = "Date", y = "AveragePrice") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "6 month") +  # Break the x-axis by months
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
        plot.title = element_text(hjust = 0.5))  # Center the title
```

Outliers Treatment

```{r}

Q1 <- quantile(avocado_data$average_price, 0.25)  # Calculate Q1 (25th percentile)
Q3 <- quantile(avocado_data$average_price, 0.75)  # Calculate Q3 (75th percentile)

# Calculate IQR
IQR <- Q3 - Q1

# Set the lower and upper limits for identifying outliers
lower_limit <- Q1 - 1.5 * IQR
upper_limit <- Q3 + 1.5 * IQR

# Filter out the outliers
avocado_data_filtered <- avocado_data %>%
  filter(average_price >= lower_limit & average_price <= upper_limit)

# View the filtered data frame
avocado_data_filtered
```

Avera

```{r}
ggplot(avocado_data_filtered, aes(x = average_price, 
                         y = factor(
                           as.character(year)
                           )
                         )
       ) +
  geom_boxplot(fill = c("#704012", "#4a7337", "#6b8c21", "#ddd48f")) +
  theme_minimal()+
  labs(x="Average Price", y="Year")
```

```{r}
ggplot(avocado_data_filtered, aes(x = average_price)) +
  geom_histogram(fill = "#4a7337",bins = 30, alpha = 0.5) +
  ggtitle("Distribution Price") +
  theme_minimal()
```

Skewness of Data

```{r}
skewness(avocado_data$average_price)
```

Skewness on a plot

```{r}
ggplot(avocado_data_filtered, aes(x = average_price)) +
  geom_histogram(fill = "#4a7337", bins = 30, alpha = 0.5) +
  geom_vline(aes(xintercept = mean(average_price)), 
             color = "blue", linetype = "dashed", size = 1) +  # Mean line
  geom_vline(aes(xintercept = median(average_price)), 
             color = "red", linetype = "dashed", size = 1) +   # Median line
  ggtitle("Distribution of Price") +
  theme_minimal()
```
If the mean is greater than the median, the distribution is right-skewed, or positively skewed and vice versa.




