---
title: "Untitled"
format: html
editor: visual
---

## **Table of Contents**

-   **Data Dictionary**

-   **Data Understanding and Loading**

-   **Exploratory Analysis Before Data Cleaning**

-   **Handling Missing Values for Numerical Variables**

-   **Handling Missing Values for Categorical Variables**

-   **Handling Missing Values with Special Characters**

-   **Outlier Treatment**

-   **Exploratory Analysis After Data Cleaning**

-   **Univariate Exploratory Analysis**

-   **Interpreting Histograms — Skewness**

-   **Interpreting Histograms — Additional Elements**

-   **Interpreting Boxplots — Quartiles and Interquartile Range**

-   **Analyzing and Interpreting Boxplots — Additional Elements**

-   **Multivariate Exploratory Analysis**

-   **Interpreting Correlation Maps**

-   **Scatter Plots — Non-Trivial Patterns**

### Libraries

```{r}
library(tidyverse)
library(janitor)
```

### Import Data

```{r}
# import data
avocado_data <- read_csv("data/avocado.csv") %>% 
  select(-1) %>% clean_names()
```

### Understanding the Structure of the Data

```{r}
str(avocado_data)
```

### Quick glimpse of data

```{r}
glimpse(avocado_data)
```

### Quick Summary Statistics

```{r}
summary(avocado_data)
```

### Missing Values

```{r}
colSums(is.na(avocado_data))
sum(is.na(avocado_data))
```
